---
title: "Hayman Fire Recovery"
author: "Tad Trimarco"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, warning=F,message=F}
library(tidyverse)
library(tidyr)
library(ggthemes)
library(lubridate)

# Now that we have learned how to munge (manipulate) data
# and plot it, we will work on using these skills in new ways

knitr::opts_knit$set(root.dir='..')
```


```{r dataread, warning=FALSE,message=FALSE}
####-----Reading in Data and Stacking it ----- ####
#Reading in files
files <- list.files('data',full.names=T)


#Read in individual data files
ndmi <- read_csv(files[1]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndmi')


ndsi <- read_csv(files[2]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndsi')

ndvi <- read_csv(files[3])%>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndvi')

# Stack as a tidy dataset
full_long <- rbind(ndvi,ndmi,ndsi) %>%
  gather(key='site',value='value',-DateTime,-data) %>%
  filter(!is.na(value))


```




## Question 1) 

What is the correlation between NDVI and NDMI? - here I want you to
convert the full_long dataset in to a wide dataset using the 
function "spread" and then make a plot that shows the correlation s a
function of if the site was burned or not (x axis should be ndmi)
You should exclude winter months and focus on summer months

```{r}

full_wide <- spread(data=full_long,key='data',value='value') %>%
  filter_if(is.numeric,all_vars(!is.na(.))) %>%
  mutate(month = month(DateTime),
         year = year(DateTime))

summer_only <- filter(full_wide,month %in% c(6,7,8,9))


ggplot(summer_only,aes(x=ndmi,y=ndvi,color=site)) + 
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.8))
```
There is a strong direct positive correlation between NDMI and NDVI. This isn't surprising, we should expect that time periods with more moisture have the greater potential to support vegetation. It is also worth noting that, unsurprisingly, NDVI is generally higher for the unburned region than the burned region, but the NDMI is also greater in the unburned region, indicating some soil moisture holding potential.

## Question 2 

2) What is the correlation between average NDSI (normalized
 snow index) for January - April and average NDVI for June-August?
In other words, does the previous year's snow cover influence vegetation
 growth for the following summer?

```{r}
jan_april <- filter(full_wide, month %in% c(1,2,3,4)) 

sum_jan <- jan_april %>%
  group_by(year,site) %>%
  summarise(mean_ndsi=mean(ndsi)) %>%
  mutate("year-site" = paste(year,site,sep=""))

sum_aug <- summer_only %>%
  group_by(year,site) %>%
  summarise(mean_ndvi = mean(ndvi)) %>%
  mutate("year-site" = paste(year,site,sep=""))

sum_2 <- inner_join(x = sum_aug, y = sum_jan, by = "year-site")

ggplot(sum_2,aes(x=mean_ndsi,y=mean_ndvi,color=site.x)) + 
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.8))
 


```
There is generally a poor correlation between 

## Q3

How is the snow effect from question 2 different between pre- and post-burn
and burned and unburned? 

```{r}
sum_3 <- sum_2 %>%
  mutate(fire = ifelse(year.x >= 2002, "Post", "Pre")) %>%
  mutate(fire_site = ifelse(site.x == "burned", paste(fire,site.x,sep="-"), site.x)) %>%
  select(-c("year-site","year.y","site.y","fire"))



ggplot(sum_3,aes(x=mean_ndsi,y=mean_ndvi,color=fire_site)) + 
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.8))
  
```


The overall lack of a significant trend is observable in both the burned and unburned plots, though there is a slightly tighter trend between NDVI and NDSI in the unburned plots, if we ignore the year (probably not wise). The only noticeable difference is that the burned plots have a wider variance in year-to-year NDVI across the same NDSI, when compared against plots that either never burn or haven't burned yet. This makes sense intuitively, extra moisture is likely to have a bigger impact on sites that actually have the vegetation to utilize it.

## Question 4

What month is the greenest month on average? 

```{r}
sum_full <- full_wide %>%
  group_by(month) %>%
  summarise(ndvi_mon =mean(ndvi),
            ndsi_mon =mean(ndsi))
greenest <- slice_max(sum_full, ndvi_mon)
greenest
```

August is the greenest on average

## Question 5) 

What month is the snowiest on average?

```{r}
snowiest <- slice_max(sum_full,ndsi_mon)
snowiest
```
January is the snowiest month on average.


## Bonus Question: Redo all problems with `spread` and `gather` using modern tidyverse syntax. 


## Bonus Question: Use Climage Engine to pull the same data for the assignment, but updated with 2020/2021 data.




