---
title: "Hayman Fire Recovery"
author: "Tad Trimarco"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, warning=F,message=F}
library(tidyverse)
library(tidyr)
library(ggthemes)
library(lubridate)

# Now that we have learned how to munge (manipulate) data
# and plot it, we will work on using these skills in new ways

knitr::opts_knit$set(root.dir='..')
```


```{r dataread, warning=FALSE,message=FALSE}
####-----Reading in Data and Stacking it ----- ####
#Reading in files
files <- list.files('data',full.names=T)


#Read in individual data files
ndmi <- read_csv(files[1]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndmi')


ndsi <- read_csv(files[2]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndsi')

ndvi <- read_csv(files[3])%>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndvi')

# Stack as a tidy dataset
full_long <- rbind(ndvi,ndmi,ndsi) %>%
  gather(key='site',value='value',-DateTime,-data) %>%
  filter(!is.na(value))


```




## Question 1) 

What is the correlation between NDVI and NDMI? - here I want you to
convert the full_long dataset in to a wide dataset using the 
function "spread" and then make a plot that shows the correlation s a
function of if the site was burned or not (x axis should be ndmi)
You should exclude winter months and focus on summer months

```{r}

full_wide <- spread(data=full_long,key='data',value='value') %>%
  filter_if(is.numeric,all_vars(!is.na(.))) %>%
  mutate(month = month(DateTime),
         year = year(DateTime))

summer_only <- filter(full_wide,month %in% c(6,7,8,9))

#Figure 1
ggplot(summer_only,aes(x=ndmi,y=ndvi,color=site)) + 
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.8))
```

Figure 1. 

There is a strong direct positive correlation between NDMI and NDVI. This isn't surprising, we should expect that time periods with more moisture have the greater potential to support vegetation. It is also worth noting that, unsurprisingly, NDVI is generally higher for the unburned region than the burned region, but the NDMI is also greater in the unburned region, indicating some soil moisture holding potential.

## Question 2 

2) What is the correlation between average NDSI (normalized
 snow index) for January - April and average NDVI for June-August?
In other words, does the previous year's snow cover influence vegetation
 growth for the following summer?

```{r}
jan_april <- filter(full_wide, month %in% c(1,2,3,4)) 

sum_jan <- jan_april %>%
  group_by(year,site) %>%
  summarise(mean_ndsi=mean(ndsi)) %>%
  mutate("year-site" = paste(year,site,sep=""))

sum_aug <- summer_only %>%
  group_by(year,site) %>%
  summarise(mean_ndvi = mean(ndvi)) %>%
  mutate("year-site" = paste(year,site,sep=""))

sum_2 <- inner_join(x = sum_aug, y = sum_jan, by = "year-site")

#Figure 2
ggplot(sum_2,aes(x=mean_ndsi,y=mean_ndvi,color=site.x)) + 
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.2))
 


```

Figure 2.

There is generally a poor correlation between NDSI and NDVI throughout the years, particularly for the burned sites. This makes sense intuitively, because while NDSI can reflect how much snow is accumulated on the ground, it happens months ahead of the productive vegetative season. A number of other factors influence how much this snow will support later vegetative growth, including snow depth, summer moisture, and whether or not the site is burned and even has vegetation to use the stored soil moisture. Notably, there does seem to be two "groupings" of burned NDSI vs NDVI values, perhaps indicating the impact of the fire on driving down NDVI values along the same NDSI value range.

## Q3

How is the snow effect from question 2 different between pre- and post-burn
and burned and unburned? 

```{r}
sum_3 <- sum_2 %>%
  mutate(fire = ifelse(year.x >= 2002, "Post", "Pre")) %>%
  mutate(fire_site = ifelse(site.x == "burned", paste(fire,site.x,sep="-"), site.x)) %>%
  select(-c("year-site","year.y","site.y","fire"))


#Figure 3
ggplot(sum_3,aes(x=mean_ndsi,y=mean_ndvi,color=fire_site)) + 
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.2))
  
```

Figure 3.

The overall lack of a significant trend is observable in both the burned and unburned plots, though there is a slightly tighter trend between NDVI and NDSI in the unburned plots, if we ignore the year (probably not wise). The only noticeable difference is that the burned plots have a wider variance in year-to-year NDVI across the same NDSI, when compared against plots that either never burn or haven't burned yet. This makes sense intuitively, extra moisture is likely to have a bigger impact on sites that actually have the vegetation to utilize it.

The correlation between NDSI and NDVI is weaker than that of NDMI and NDVI. Intuitively, this makes sense. NDMI values from Figure 2 are from the same time period as the NDVI values, whereas the NDSI values in Figure 3 come from the earlier winter months and more represent the "potential" for moisture that can be stored in soil and used later. NDMI values are a representation of the actual moisture available to the plants during the growing season, so we should expect to see a stronger correlation between NDMI and NDVI here.

## Question 4

What month is the greenest month on average? 

```{r}
sum_full <- full_wide %>%
  group_by(month) %>%
  summarise(ndvi_mon =mean(ndvi),
            ndsi_mon =mean(ndsi))
greenest <- slice_max(sum_full, ndvi_mon)
greenest
```

August is the greenest on average.

## Question 5) 

What month is the snowiest on average?

```{r}
snowiest <- slice_max(sum_full,ndsi_mon)
snowiest
```
January is the snowiest month on average.






